Title Goes Here

### 2017-10-26 analysis/paper/saa2018_general_sessions.Rmd

```{r eval=FALSE}
# read in the abstracts
general <- read_excel("../data/raw_data/General Abstracts.xlsx")
```

@@

```{r eval=FALSE}
# Create a corpus object from the abstracts
general_corpus <- tm::Corpus(tm::DataframeSource(general), 
                             readerControl = list(reader = tm::readTabular(mapping = m)))

# create a document term matrix
general_tm <- tm::DocumentTermMatrix(general_corpus, 
                                     control = list(stemming = TRUE, 
                                                    stopwords = TRUE,
                                                    minWordLength = 2, 
                      s                              removeNumbers = TRUE, 
                                                    removePunctuation = TRUE))
```

```{r eval=FALSE}
# Use TF-IDF to wieght terms and remove rare ones
term_tfidf <- 
  tapply(general_tm$v/slam::row_sums(general_tm)[general_tm$i], 
         general_tm$j, mean) * log2(tm::nDocs(general_tm)/slam::col_sums(general_tm > 0))

summary(term_tfidf)
# Median =  0.08989
```

```{r eval=FALSE }
## Keeping the rows with tfidf >= median
general_reduced_dtm <- general_tm[,term_tfidf >= summary(term_tfidf)[3]]
summary(slam::col_sums(general_reduced_dtm))
# save
saveRDS(general_reduced_dtm, "../data/derived_data/general_reduced_dtm.rds")
```

```{r}
general_reduced_dtm <- readRDS("../data/derived_data/general_reduced_dtm.rds")
```

@@

```{r}
# run a loop over the abstract with different numbers of topics to find which number of topcis is the best
seqk <- seq(2, 100, 1)
burnin <- 1000
iter <- 1000
keep <- 50
```

```{r eval=FALSE }
system.time(fitted_many <- 
              lapply(seqk, function(k) topicmodels::LDA(general_reduced_dtm, 
                                                        k = k,
                                                        method = "Gibbs",
                                                        control = list(burnin = burnin,
                                                                       iter = iter, 
                                                                       keep = keep) )))

saveRDS(fitted_many, "../data/derived_data/fitted_many.rds")
```

```{r}
fitted_many <- readRDS("../data/derived_data/fitted_many.rds")
```

```{r}
# extract logliks from each topic
logLiks_many <- lapply(fitted_many, function(L)  L@logLiks[-c(1:(burnin/keep))])

# compute harmonic means
hm_many <- sapply(logLiks_many, function(h) harmonicMean(h))

# to know the optimal number of topics:
optimal_number_of_topics <- seqk[which.max(hm_many)]

# plot it
library(ggplot2)
lda_plot <- ggplot(data.frame(seqk, 
                             hm_many), 
                  aes(x=seqk, 
                      y=hm_many)) + 
  geom_path(lwd=1.5)  +
  xlab('Number of Topics') +
  ylab('Harmonic Mean') +
  geom_vline(xintercept = seqk[which.max(hm_many)],
             colour = "red") +
  annotate("text",
           x = 55, 
           y = -150000, 
           label = paste("The optimal number of topics is", 
                         seqk[which.max(hm_many)])) +
  theme_bw() +
  ggtitle(expression(atop("Latent Dirichlet Allocation Analysis of SAA General Abstracts", atop(italic("How many distinct topics in the abstracts?"), ""))))
```

```{r}
# now run the model with the optimum number of topics
system.time(general_model <- topicmodels::LDA(general_reduced_dtm, 
                                              optimal_number_of_topics,
                                              method = "Gibbs", 
                                              control = list(iter=2000, 
                                                             seed = 0622)))

# explore the model
general_topics <- topicmodels::topics(general_model, 1)

## In this case I am returning the top 30 terms.
general_terms <- as.data.frame(topicmodels::terms(general_model, 30), 
                               stringsAsFactors = FALSE)
```

```{r}
library(tidytext)
tidy_topics <- tidy(general_model, matrix = "beta")

@@

tidy_topics_top_terms <- tidy_topics %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)
```

```{r}
tidy_topics_top_terms %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()
```

```{r}
tidy_documents <- tidy(general_model, matrix = "gamma")

library(tidyr)
tidy_documents_topics_gamma <-
  tidy_documents %>% 
    spread(topic, gamma)

abstract_classifications <- 
  tidy_documents %>%
  group_by(document) %>%
  top_n(1, gamma) %>%
  ungroup() %>% 
  arrange(desc(gamma))

abstract_classifications
```

>>>>

```{r eval=TRUE}
# read in the abstracts
general_2018 <- read_excel("../data/raw_data/General Abstracts_2018.xlsx")

general_2018_posters <- general_2018 %>% filter(grepl("Poster", `Presentation Format`))
general_2018_papers  <- general_2018 %>% filter(grepl("Paper", `Presentation Format`))
```

@@

```{r eval=FALSE}
# Create a corpus object from the abstracts
general_2018_papers_corpus <- tm::Corpus(tm::DataframeSource(general_2018_papers), 
                             readerControl = list(reader = tm::readTabular(mapping = m)))

# create a document term matrix
general_2018_papers_tm <- tm::DocumentTermMatrix(general_2018_papers_corpus, 
                                     control = list(stemming = TRUE, 
                                                    stopwords = TRUE,
                                                    minWordLength = 2, 
                                                    removeNumbers = TRUE, 
                                                    removePunctuation = TRUE))

```

```{r eval=FALSE}
# Use TF-IDF to wieght terms and remove rare ones
term_2018_papers_tfidf <- 
  tapply(general_2018_papers_tm$v/slam::row_sums(general_2018_papers_tm)[general_2018_papers_tm$i], 
         general_2018_papers_tm$j, mean) * log2(tm::nDocs(general_2018_papers_tm)/slam::col_sums(general_2018_papers_tm > 0))

summary(term_2018_papers_tfidf)
# Median =   0.08120
```

```{r eval=FALSE }
## Keeping the rows with tfidf >= median
general_2018_papers_reduced_dtm <- general_2018_papers_tm[,term_2018_papers_tfidf >= summary(term_2018_papers_tfidf)[3]]
summary(slam::col_sums(general_2018_papers_reduced_dtm))
# save
saveRDS(general_2018_papers_reduced_dtm, "../data/derived_data/general_2018_papers_reduced_dtmrds")
```

```{r}
general_2018_papers_reduced_dtm <- readRDS("../data/derived_data/general_2018_papers_reduced_dtm.rds")
number_of_paper_abstracts <- general_2018_papers_reduced_dtm$nrow
```

@@

```{r}
# run a loop over the abstract with different numbers of topics to find which number of topics is the best
seqk <- seq(2, 100, 1)
burnin <- 1000
iter <- 1000
keep <- 50
```

```{r eval=FALSE }
system.time(fitted_many_2018_papers <- 
              lapply(seqk, function(k) topicmodels::LDA(general_2018_papers_reduced_dtm, 
                                                        k = k,
                                                        method = "Gibbs",
                                                        control = list(burnin = burnin,
                                                                       iter = iter, 
                                                                       keep = keep) )))

saveRDS(fitted_many_2018_papers, "../data/derived_data/fitted_many_2018_papers.rds")

# rerun with k = 60 to force the abstracts into 60 topics
```

```{r}
fitted_many_2018_papers <- readRDS("../data/derived_data/fitted_many_2018_papers.rds")

# how many abstracts, if we just start from here?
how_many_paper_abstracts <- fitted_many_2018_papers[[1]]@Dim[1]
```

```{r}
# extract logliks from each topic
logLiks_many_2018_papers <- lapply(fitted_many_2018_papers, function(L)  L@logLiks[-c(1:(burnin/keep))])

# compute harmonic means
hm_many_2018_papers <- sapply(logLiks_many_2018_papers, function(h) harmonicMean(h))

# to know the optimal number of topics:
optimal_number_of_topics_2018_papers <- seqk[which.max(hm_many_2018_papers)]

# plot it
library(ggplot2)
lda_plot <- ggplot(data.frame(seqk, 
                             hm_many_2018_papers), 
                  aes(x=seqk, 
                      y=hm_many_2018_papers)) + 
  geom_path(lwd=1.5)  +
  xlab('Number of Topics') +
  ylab('Harmonic Mean') +
  geom_vline(xintercept = optimal_number_of_topics_2018_papers,
             colour = "red") +
  annotate("text",
           x = 55, 
           y = -150000, 
           label = paste("The optimal number of topics is", 
                         optimal_number_of_topics_2018_papers)) +
  theme_bw() +
  ggtitle("Latent Dirichlet Allocation Analysis of SAA General Abstracts",
          subtitle = paste0("How many distinct topics are present in the general paper abstracts (n = ", how_many_paper_abstracts, ") for 2018?"))

ggsave("../figures/saa2018_general_abstracts_papers_opti_topics.png")
```

```{r}
# now run the model with the optimum number of topics
system.time(general_model_papers <- topicmodels::LDA(general_2018_papers_reduced_dtm, 
                                              optimal_number_of_topics_2018_papers,
                                              method = "Gibbs", 
                                              control = list(iter=2000, 
                                                             seed = 0622)))

# explore the model
general_topics_papers <- topicmodels::topics(general_model_papers, 1)

## In this case I am returning the top 30 terms.
general_terms_papers <- as.data.frame(topicmodels::terms(general_model_papers, 30), 
                               stringsAsFactors = FALSE)
```

```{r}
library(tidytext)
tidy_topics_papers <- tidy(general_model_papers, matrix = "beta")

@@

tidy_topics_papers_top_terms <- tidy_topics_papers %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)
```

```{r}
# show top keywords for each topic
tidy_topics_papers_top_terms %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()

ggsave("../figures/keywords_for_saa2018_general_papers_topics.png", h = 14, w = 14)
```

```{r}
tidy_documents_papers <- tidy(general_model_papers, matrix = "gamma")

library(tidyr)
tidy_documents_papers_topics_gamma <-
  tidy_documents_papers %>% 
    spread(topic, gamma)

abstract_classifications_papers <- 
  tidy_documents_papers %>%
  group_by(document) %>%
  top_n(1, gamma) %>%
  ungroup() %>% 
  arrange(desc(gamma)) %>% 
  mutate(document = as.numeric(document))

abstract_classifications_papers
```

```{r}
# put topics back onto abstract spreadsheet
# abstract_ids <- dimnames(general_2018_reduced_dtm)$Docs
# general_2018$`Abstract Id`

general_2018_papers_with_topics <- 
  general_2018_papers %>% 
  left_join(abstract_classifications_papers, 
            by = c("Abstract Id" = "document"))

write.csv(general_2018_papers_with_topics, "../data/derived_data/general_2018_papers_with_topics.csv")
```

### 2017-10-26 analysis/paper/saa2018_general_sessions_hclust.Rmd

# clean geo
clean_geo <- function(x){
  tmp <- stringr::str_trim(tolower(gsub("[[:punct:]]|\r\n| and ", " ", x)))
  tmp <- gsub(" ", "_", tmp)
  tmp <- gsub("_{2}", "_", tmp)
  tmp
}

@@

ndocs <- length(general_2018_papers_corpus)
# ignore extremely rare words i.e. terms that appear in less then 1% of the documents
minTermFreq <- ndocs * 0.01
# ignore overly common words i.e. terms that appear in more than 50% of the documents
maxTermFreq <- ndocs * .5

@@

```{r}
m  <- as.matrix(general_2018_papers_tm)
distMatrix <- dist(m, method="euclidean")

groups <- hclust(distMatrix,method="ward.D")
plot(groups, cex=0.9, hang=-1)
rect.hclust(groups, k=70)

# see what cluster each abstract is in:
cuts <- cutree(groups, k = 50)
```

>>>>

# clean geo
clean_geo <- function(x){
  tmp <- stringr::str_trim(tolower(gsub("[[:punct:]]|\r\n| and ", " ", x)))
  tmp <- gsub("other", "", tmp)
  tmp <- gsub(" ", "_", tmp)
  tmp <- gsub("_{2}", "_", tmp)
  tmp
}

@@

ndocs <- length(general_2018_papers_corpus)
# ignore extremely rare words i.e. terms that appear in less then 1% of the documents
minTermFreq <- ndocs * 0.01
# ignore overly common words i.e. terms that appear in more than 50% of the documents
maxTermFreq <- ndocs * 0.5

@@

Distance matrix

```{r}
m  <- as.matrix(general_2018_papers_tm)
distMatrix <- dist(m, method="euclidean")
```

How many clusters?

```{r}
library(apcluster)
d.apclus <- apcluster(negDistMat(r=2), m)
k <- length(d.apclus@clusters)

windows()
heatmap(d.apclus)
```

Cluster analysis and dendrogram

```{r}
groups <- hclust(distMatrix,method="ward.D")
plot(groups, cex=0.9, hang=-1)
rect.hclust(groups, k)

# see what cluster each abstract is in:
cuts <- cutree(groups, k)
```

How many items in each cluster

```{r}
# histogram
freqs_per_cluster <- 
cuts %>%
  data_frame(cuts = .,
             freq = names(.)) %>% 
  group_by(cuts) %>% 
  tally()

hist(freqs_per_cluster$n)

median(freqs_per_cluster$n) # 10
```

### 2017-10-26 analysis/paper/saa2018_general_sessions_papers_hclust.Rmd

```{r eval=TRUE}
# read in the abstracts
general_2018 <- read_excel("../data/raw_data/General Abstracts_2018.xlsx")

@@

general_2018_papers$text_to_model <- 
  with(general_2018_papers, paste(geo, kw1, kw2, kw3))
```

@@

```{r}
# histogram
freqs_per_cluster <- 
cuts %>%
  data_frame(cuts = .,
             freq = names(.)) %>% 
  group_by(cuts) %>% 
  tally()

@@

>>>>

```{r eval=TRUE}
# read in the abstracts
general_2018 <- ab# read_excel("../data/raw_data/General Abstracts_2018.xlsx")

@@

library(stringi)
general_2018_papers$text_to_model <- 
  paste(stri_dup(paste(general_2018_papers$geo, ""), 4),
        stri_dup(paste(general_2018_papers$kw1, ""), 3),
        general_2018_papers$kw2, 
        general_2018_papers$kw3)
```

@@

```{r}
# histogram
freqs_per_cluster <- 
cuts %>%
  data_frame(cuts = .,
             freq = names(.)) %>% 
  group_by(cuts) %>% 
  tally(sort = TRUE)

@@

Compare

```{r}
ab <- read_excel("../data/derived_data/Copy of general_2018_papers_with_clusters_AB.xlsx")

compare <- left_join(general_2018_papers_with_topics, ab, by = "Abstract Id")

compare_groups <- 
compare %>% 
  select(cuts.x, cuts.y)
```

reclassify using weighted keywords

```{r}
